---
title: "HOMEWORK 2"
output: 
  html_document: 
    css: ~/git/CMPE/fall19-ceyonur/docs/hacker.css
    df_print: paged
---
## **Ceyhun Onur**

## _01/11/2019_

# Task 1

## Part 1
```{r eval=FALSE, include=FALSE}
install.packages("data.table")
install.packages("factoextra")
```

```{r message=FALSE, warning=FALSE , echo=T, results="hide"}
library(data.table)
library(factoextra)
```

```{r}
musk <- fread("/home/ceyonur/git/CMPE/fall19-ceyonur/data/homework2/Musk1.csv")
# rename first two columns
names(musk)[1:2]<-c("bag_class", "bag_id")
musk
```

```{r}
summary_musk_table <- data.frame(unclass(summary(musk)), check.names = FALSE, stringsAsFactors = FALSE)
summary_musk_table
```

### PCA
```{r}
# remove first two columns (bag class and bag id)
raw_musk <- musk[,c(-1,-2)]
pca_musk <- prcomp(raw_musk)
fviz_eig(pca_musk, addlabels = TRUE)
summary(pca_musk)
```

```{r}
fviz_pca_ind(pca_musk, habillage = musk$bag_class, palette = c("red", "black"), addEllipses = TRUE)
```

**It seems like PCA values are overlapping. So we cannot derive any "useful" grouping information from here.**

### MDS
```{r}
eu_matrix <- dist(raw_musk, method="euclidean") 
eu_matrix[is.na(eu_matrix)] <- 0

man_matrix <- dist(raw_musk, method="manhattan") 
man_matrix[is.na(man_matrix)] <- 0
```

```{r}
mds_eu <- cmdscale(eu_matrix)
plot(mds_eu[,1],mds_eu[,2],main='MDS Euclidean',xlab='', ylab='',col=musk$bag_class + 1)
```

```{r}
mds_man <- cmdscale(man_matrix)
plot(mds_man[,1],mds_man[,2],main='MDS Manhattan',xlab='', ylab='',col=musk$bag_class + 1 )
```

**MDS results are also similar to PCA results, both of them does not seperate the results. They both are overlapping.**

## Part 2 - Summarizing data

```{r}
reduced_musk <- musk[,lapply(.SD, mean), by=bag_id]
reduced_musk
```

### PCA
```{r}
raw_reduced_musk <- reduced_musk[,c(-1,-2)]
pca_reduced_musk <- prcomp(raw_reduced_musk)
fviz_eig(pca_reduced_musk, addlabels = TRUE)
summary(pca_reduced_musk)
```

```{r}
fviz_pca_ind(pca_reduced_musk, habillage = reduced_musk$bag_class, palette = c("red", "black"),addEllipses = TRUE)
```

**When compared to the first PCA plot, this one is less overlapping, the group 0 has a much seperate group. However it's still overlapping a lot. So even with the grouped results, we cannot say PCA clearly tells us how to group values.**

## MDS
```{r}
eu_matrix_r <- dist(reduced_musk, method="euclidean") 
eu_matrix_r[is.na(eu_matrix_r)] <- 0

man_matrix_r <- dist(reduced_musk, method="manhattan") 
man_matrix_r[is.na(man_matrix_r)] <- 0

mds_eu_r <- cmdscale(eu_matrix_r)
plot(mds_eu_r[,1],mds_eu_r[,2],main='MDS Euclidean',xlab='', ylab='',col=reduced_musk$bag_class + 1)
```
```{r}
mds_man_r=cmdscale(man_matrix_r)
plot(mds_man_r[,1],mds_man_r[,2],main='MDS Manhattan',xlab='', ylab='',col=reduced_musk$bag_class + 1 )
```

**MDS plots also cannot tell us more when they're grouped.**

# Task 2

## Part 1 - Read and Show the Image

```{r eval=FALSE, include=FALSE}
install.packages("jpeg")
```

```{r message=FALSE, warning=FALSE , echo=T, results="hide"}
library(jpeg)
```

```{r}

lake <- readJPEG("/home/ceyonur/git/CMPE/fall19-ceyonur/data/homework2/IMG-20190421-WA0028.jpg")
# dimensions of the JPEG
dim(lake)
```

```{r}
plot(0:256, 0:256, xlab= "x",ylab="y", main="lake", frame.plot = TRUE, axes= FALSE, type="n")
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
axy<-axis(2, at = yticks, labels = yticks)
axx<-axis(1, at = xticks, labels = xticks)
rasterImage(lake, 0, 0, 256, 256,interpolate=FALSE)
```

## Part 2 - Noisy Image
### Subpart a - Show Image

```{r}
reds <- lake[,,1]
greens <- lake[,,2]
blues <- lake[,,3]
rednoise <- matrix(runif(256*256, min(reds),max(reds)*0.1), 256,256)
greennoise <- matrix(runif(256*256, min(greens),max(greens)*0.1), 256,256)
bluenoise <- matrix(runif(256*256, min(blues),max(blues)*0.1), 256,256)
noisy_lake <- lake
noisy_lake[,,1] <- reds + rednoise
noisy_lake[,,2] <- greens + greennoise
noisy_lake[,,3] <- blues + bluenoise
noisy_lake<-ifelse(noisy_lake>1,1,noisy_lake)

plot(0:256, 0:256, xlab= "x",ylab="y", main="noisy lake", frame.plot = TRUE, axes= FALSE, type="n")
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(noisy_lake, 0, 0, 256, 256,interpolate=FALSE)
```
```{r}
summary(noisy_lake - lake)
```

```{r}
plot(0:256, 0:256, xlab= "x",ylab="y", main="noise", frame.plot = TRUE, axes= FALSE, type="n")
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(noisy_lake - lake, 0, 0, 256, 256,interpolate=FALSE)
```

### Subpart b - Color Channels
```{r}
par(mfrow=c(1,3), oma = c(0, 0, 2, 0))

image(1:256,1:256,t(apply(lake[,,1],2,rev)), xlab="x",ylab="y", main="red channel")
image(1:256,1:256,t(apply(lake[,,2],2,rev)), xlab="x",ylab="y", main="green channel")
image(1:256,1:256,t(apply(lake[,,3],2,rev)), xlab="x",ylab="y", main="blue channel")
mtext("Original RGB Channels", outer = TRUE, cex = 1.5)

par(mfrow=c(1,3), oma = c(0, 0, 2, 0))
image(1:256,1:256,t(apply(noisy_lake[,,1],2,rev)), xlab="x",ylab="y", main="red channel")
image(1:256,1:256,t(apply(noisy_lake[,,2],2,rev)), xlab="x",ylab="y", main="green channel")
image(1:256,1:256,t(apply(noisy_lake[,,3],2,rev)), xlab="x",ylab="y", main="blue channel")
mtext("Noisy RGB Channels", outer = TRUE, cex = 1.5)
```

```{r}
par(mfrow=c(1,3), oma = c(0, 0, 2, 0))
lake_red <- lake
lake_red[,,2:3] <- 0
lake_blue <- lake
lake_blue[,,c(1,3)] <- 0
lake_green <- lake
lake_green[,,1:2] <- 0
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
plot(0:256, 0:256, xlab = "x",ylab = "y", main = "red channel", frame.plot = TRUE, axes = FALSE, type = "n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(lake_red, 0, 0, 256, 256, interpolate = FALSE)

plot(0:256, 0:256, xlab = "x",ylab = "y", main = "blue channel", frame.plot = TRUE, axes = FALSE, type = "n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(lake_blue, 0, 0, 256, 256, interpolate = FALSE)

plot(0:256, 0:256, xlab = "x", ylab = "y", main = "green channel", frame.plot = TRUE, axes = FALSE, type = "n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(lake_green, 0, 0, 256, 256, interpolate = FALSE)
mtext("Original RGB Channels", outer = TRUE, cex = 1.5)

par(mfrow = c(1,3), oma = c(0, 0, 2, 0))
noisy_lake_red <- noisy_lake
noisy_lake_red[,,2:3] <- 0
noisy_lake_blue <- noisy_lake
noisy_lake_blue[,,c(1,3)] <- 0
noisy_lake_green <- noisy_lake
noisy_lake_green[,,1:2] <- 0
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
plot(0:256, 0:256, xlab = "x",ylab = "y", main ="red channel", frame.plot = TRUE, axes= FALSE, type="n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(noisy_lake_red, 0, 0, 256, 256,interpolate=FALSE)

plot(0:256, 0:256, xlab= "x",ylab="y", main="blue channel", frame.plot = TRUE, axes= FALSE, type="n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(noisy_lake_blue, 0, 0, 256, 256,interpolate=FALSE)

plot(0:256, 0:256, xlab= "x",ylab="y", main="green channel", frame.plot = TRUE, axes= FALSE, type="n")
axis(2, at = yticks, labels = yticks)
axis(1, at = xticks, labels = xticks)
rasterImage(noisy_lake_green, 0, 0, 256, 256,interpolate=FALSE)
mtext("Noisy RGB Channels", outer = TRUE, cex = 1.5)
```

```{r}
par(mfrow=c(1,2))
plot(lake[,,1], col="red", main = "Original", xlab="", ylab="")
points(lake[,,2], col="green")
points(lake[,,3], col="blue")

plot(noisy_lake[,,1], col="red", main = "Noisy", xlab="", ylab="")
points(noisy_lake[,,2], col="green")
points(noisy_lake[,,3], col="blue")
mtext("Color Intensities", outer = TRUE, cex = 1.5)
```

## Part 3 - PCA

```{r}
gray_noisy_lake<- noisy_lake[,,1] + noisy_lake[,,2] + noisy_lake[,,3]
# normalize
gray_noisy_lake <- gray_noisy_lake/max(gray_noisy_lake)
plot(0:256, 0:256, xlab= "x",ylab="y", main="gray noisy lake", frame.plot = TRUE, axes= FALSE, type="n")
xticks <-seq(0,256,256/4)
yticks <-seq(0,256,256/4)
axy<-axis(2, at = yticks, labels = yticks)
axx<-axis(1, at = xticks, labels = xticks)
rasterImage(gray_noisy_lake, 0, 0, 256, 256,interpolate = FALSE)
```

```{r}
patch_size <- 3
from <- floor(patch_size/2) + 1
to <- dim(gray_noisy_lake)[1] - from + 1
patches <- matrix(0,nrow = length(from:to)^2,ncol=patch_size^2, byrow=TRUE)
margin <- floor(patch_size/2)
counter <- 1
for(y in from:to){
  for(x in from:to){
    patch <- gray_noisy_lake[(x - margin):(x + margin),(y - margin):(y + margin) ]
    patches[counter, ] <- as.vector(patch)
    counter <- counter + 1
  }
}
gray_noisy_lake_pca<-prcomp(patches)
plot(gray_noisy_lake_pca)
```
```{r}
summary(gray_noisy_lake_pca)
```

```{r}
scores <- gray_noisy_lake_pca$x
comp1_matrix <- matrix(scores[,1], nrow = 254, ncol = 254 , byrow = TRUE)
plot(c(0, 254), c(0, 254), type = "n", xlab = "", ylab = "",main = "Comp.1 Image")

normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x)))
}
comp1_matrix <- t(comp1_matrix)

rasterImage(normalize(comp1_matrix),0,0,254,254)
```

```{r}
comp2_matrix <- matrix(scores[,2], nrow = 254, ncol = 254 , byrow = TRUE)
plot(c(0, 254), c(0, 254), type = "n", xlab = "", ylab = "",main = "Comp.2 Image")

comp2_matrix <- t(comp2_matrix)

rasterImage(normalize(comp2_matrix),0,0,254,254)
```

```{r}
comp3_matrix <- matrix(scores[,3], nrow = 254, ncol = 254 , byrow = TRUE)
plot(c(0, 254), c(0, 254), type = "n", xlab = "", ylab = "",main = "Comp.3 Image")

comp3_matrix <- t(comp3_matrix)

rasterImage(normalize(comp3_matrix),0,0,254,254)
```

```{r}
eigenVectors <- gray_noisy_lake_pca$rotation[,1:3]
eigenVectors
```

```{r}
comp1_eigens <- matrix(eigenVectors[,1], nrow = patch_size, ncol = patch_size , byrow = TRUE)
plot(c(0, patch_size), c(0, patch_size), type = "n", xlab = "", ylab = "",main = "Eigen Comp.1 Image")

comp1_eigens <- t(comp1_eigens)

rasterImage(normalize(comp1_eigens),0,0,patch_size,patch_size, interpolate = FALSE)
```

```{r}
comp2_eigens <- matrix(eigenVectors[,2], nrow = patch_size, ncol = patch_size , byrow = TRUE)
plot(c(0, patch_size), c(0, patch_size), type = "n", xlab = "", ylab = "",main = "Eigen Comp.2 Image")

comp2_eigens <- t(comp2_eigens)

rasterImage(normalize(comp2_eigens),0,0,patch_size,patch_size, interpolate = FALSE)
```

```{r}
comp3_eigens <- matrix(eigenVectors[,3], nrow = patch_size, ncol = patch_size , byrow = TRUE)
plot(c(0, patch_size), c(0, patch_size), type = "n", xlab = "", ylab = "",main = "Eigen Comp.3 Image")

comp3_eigens <- t(comp3_eigens)

rasterImage(normalize(comp3_eigens),0,0,patch_size,patch_size, interpolate = FALSE)
```